---
title: "9.26.2019"
output: github_document
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(ggridges)
##
##Attaching package: 'ggridges'
## The following object is masked from 'package:ggplot2':
##
##    scale_discrete_manual
```

# Weather Data
```{r}
weather_df = 
  rnoaa::meteo_pull_monitors(c("USW00094728", "USC00519397", "USS0023B17S"),
                      var = c("PRCP", "TMIN", "TMAX"), 
                      date_min = "2017-01-01",
                      date_max = "2017-12-31") %>%
  mutate(
    name = recode(id, USW00094728 = "CentralPark_NY", 
                      USC00519397 = "Waikiki_HA",
                      USS0023B17S = "Waterhole_WA"),
    tmin = tmin / 10,
    tmax = tmax / 10) %>%
  select(name, id, everything())
weather_df
##
## Create a ggplot, aes= aesthetic mapping we care about, without geompoint() we don't have anything on our graph
## geompoint() at a point at each x/y variable we will put a point
##
ggplot(weather_df, aes(x = tmin, y = tmax)) + 
  geom_point()
##
## 
```
# Different aesthetic mapping
```{r}
weather_df %>%
  ggplot(aes(x = tmin, y = tmax)) +
  geom_point()
##
## Another way of making above ggplot

```

## Filtering by specific variable name
```{r}
weather_df %>% filter(name == "Central_Park_NY")
```

# saving plot with a name (plot_weather)
```{r}
plot_weather = 
  weather_df %>%
  ggplot(aes(x = tmin, y = tmax))

plot_weather + geom_point()
```

Adding color... 

```{r}
ggplot(weather_df, aes(x = tmin, y = tmax)) + 
  geom_point(aes(color = name), alpha = 0.4)
##
## Adding color into the different points, each color represents a different name (weather station location)
## Alpha = 0.4 adds transparency to data points
```

## Adding line through the data (smooth, curvy best line), se= FALSE turns error bars off
* First
```{r}
ggplot(weather_df, aes(x = tmin, y = tmax)) + 
  geom_point(aes(color = name), alpha = .5) +
  geom_smooth(se = FALSE)
```
* vs, not defining all aesthetic mappings for everything (all three variables), color matters in geom_point but not in geom_smooth
```{r}
ggplot(weather_df, aes(x = tmin, y = tmax, color = name)) + 
  geom_point(alpha = .5) +
  geom_smooth(se = FALSE)
##
## Gives you smooth lines for each variable (one for each location of weather temps)
```

## Want to turn one panel plot into three panel plot, one plot for each weather station
```{r}
ggplot(weather_df, aes(x = tmin, y = tmax, color = name)) + 
  geom_point(alpha = .5) +
  geom_smooth(se = FALSE) + 
  facet_grid(. ~ name)
##
## Facets break out your plot by the variable you specify (here, you break out by name)
## Easier to see what's going on in each different weather station, helpful to show individual variables
```

## the points are sized based on how much precipitation happens on that date, temperature based on day of the year (so big curves in WA and NY because seasons, Waikiki is fairly similar temp all year round)
```{r}
ggplot(weather_df, aes(x = date, y = tmax, color = name)) + 
  geom_point(aes(size = prcp), alpha = .5) +
  geom_smooth(se = FALSE) + 
  facet_grid(. ~ name)
##
## Color= name so color by different weather stations
## and break out by facet to show each variable, can do all together to compare on top of each other
##
```
# shows precipitation based on day, not as good as above
```{r}
ggplot(weather_df, aes(x = date, y = prcp, color = name)) + 
  geom_point( alpha = .5) +
  geom_smooth(se = FALSE) 
```
```{r}
weather_df %>% 
  filter(name == "CentralPark_NY") %>% 
  mutate(tmax_fahr = tmax * (9 / 5) + 32,
         tmin_fahr = tmin * (9 / 5) + 32) %>% 
  ggplot(aes(x = tmin_fahr, y = tmax_fahr)) +
  geom_point(alpha = .5) + 
  geom_smooth(method = "lm", se = FALSE)
## Only central park, converting to celsius
```

## Some extra stuff
```{r}
ggplot(weather_df, aes(x = date, y = tmax, color = name)) + 
  geom_smooth(se = FALSE) 
##
## If take out geom_point then you get rid of individual points, can do geom_smooth to just show smooth curves that show average t_max
## inferior to plots that show data points
```
### Different kinds of geoms
```{r}
weather_df %>%
  ggplot(aes(x = tmin, y = tmax)) + 
  geom_bin2d()
##
## Square geoms
```

# Univariate Plots
```{r}
ggplot(weather_df, aes(x = tmax)) + 
  geom_histogram()
##
## Distribution of only tmax as histogram
```

```{r}
ggplot(weather_df, aes(x = tmax, fill = name)) + 
  geom_histogram(position = "dodge", binwidth = 2)
##
## puts color into the bars that correspond with weather station given tmax
## position makes the bars next to each other and "dodge" puts the bars not on top (dodge each other)
## for histogram: "color" shows outside color, and "fill" is coloring the inside of the bar
```
```{r}
ggplot(weather_df, aes(x = tmax, fill = name)) + 
  geom_histogram(position = "dodge", binwidth = 2) +
  facet_grid(~name)
##
## three separate histograms
```


```{r}
ggplot(weather_df, aes(x = tmax, fill = name)) + 
  geom_density(alpha = .4, adjust = .5, color = "blue")
## 
## Density curves are better way to show how many of each tmax ther are in each weather station
## alpha = 0.3 makes it easier to see, color makes the outline blue to have a better idea of top
## 
```

## Boxplots
```{r}
ggplot(weather_df, aes(x = name, y = tmax)) + geom_boxplot()
##
## for each location, you get a spread of the t max's
##
```

## Violing plot
```{r}
weather_df %>%
ggplot(aes(x = name, y = tmax)) + 
  geom_violin(aes(fill = name), color = "blue", alpha = .5) + 
  stat_summary(fun.y = median, geom = "point", color = "blue", size = 4)
##
## Density plot version of a box plot (same kind of density plot info but just vertical)
```

## Ridge plots
```{r}
ggplot(weather_df, aes(x = tmax, y = name)) + 
  geom_density_ridges(scale = .85)
##
## Shows a density of tmax across time horizontally
```
## saving a plot
```{r}
weather_plot = ggplot(weather_df, aes(x = tmin, y = tmax)) + 
  geom_point(aes(color = name), alpha = .5) 

ggsave("weather_plot.pdf", weather_plot, width = 8, height = 5)
##
## now it's saved in my folder as a pdf on my computer 
```

```{r, fig. width = 12}
weather_df %>%
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = .4) +
  geom_smooth(se = FALSE)
```
```{r, fig. width = 6}
weather_df %>%
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = .4) +
  geom_smooth(se = FALSE)
```
 
## 10/1/2019 Notes
```{r}
weather_df %>%
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = .4)
```

## Adding labels to the ggplot
```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax)) + 
  geom_point(aes(color = name), alpha = .5) + 
  labs(
    title = "Temperature plot",
    x = "Minimum daily temperature (C)",
    y = "Maxiumum daily temperature (C)",
    caption = "Data from the rnoaa package"
  )
```
## Playing with the scales
```{r} 
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax)) + 
  geom_point(aes(color = name), alpha = .5) + 
  labs(
    title = "Temperature plot",
    x = "Minimum daily temperature (C)",
    y = "Maxiumum daily temperature (C)",
    caption = "Data from the rnoaa package") + 
  scale_x_continuous(
    breaks = c(-15, 0, 15), 
    labels = c("-15ยบ C", "0", "15"))

## This scales your axes 
```
# Transformation
```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax)) + 
  geom_point(aes(color = name), alpha = .5) + 
  labs(
    title = "Temperature plot",
    x = "Minimum daily temperature (C)",
    y = "Maxiumum daily temperature (C)",
    caption = "Data from the rnoaa package") + 
  scale_x_continuous(
    breaks = c(-15, 0, 15), 
    labels = c("-15ยบ C", "0", "15"),
    limits = c(-20, 30)
)+ 
  scale_y_continuous(
    trans = "sqrt", 
    position = "right")
##Transforms your data with a squareroot to your scale
```
#Colors
```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax)) + 
  geom_point(aes(color = name), alpha = .5) + 
  labs(
    title = "Temperature plot",
    x = "Minimum daily temperature (C)",
    y = "Maxiumum daily temperature (C)",
    caption = "Data from the rnoaa package") + 
  scale_color_hue(name = "Location", h = c(100, 300))

## scale_color_hue (many different options, can play around, not recommended)
```

## Viridis package, solves very specific problem in tidyverse
* ggplot colors aren't great so this solves this (better color palates)
* more distinct colors, light to dark
```{r}
ggp_base = 
  weather_df %>% 
  ggplot(aes(x = tmin, y = tmax)) + 
  geom_point(aes(color = name), alpha = .5) + 
  labs(
    title = "Temperature plot",
    x = "Minimum daily temperature (C)",
    y = "Maxiumum daily temperature (C)",
    caption = "Data from the rnoaa package"
  ) + 
  viridis::scale_color_viridis(
    name = "Location", 
    discrete = TRUE
  )

ggp_base

## Discrete = TRUE bc variable is discrete
```
#changing other fun things to ggp_base
#scales = how axes demonstrate specific mapping, now we do themes, global things that describe what plot looks like, general behavior and look of plot
```{r}
#legend position
ggp_base +
  theme_bw() +
  theme(legend.position = "bottom") 

  #take legend out completely = legend.position = "none"
  #Order really matters
#theme_minimal() makes it all minimal
```
# changing geom line to connect the dots
```{r}
central_park = 
  weather_df %>% 
  filter(name == "CentralPark_NY")

waikiki = 
  weather_df %>% 
  filter(name == "Waikiki_HA")

ggplot(data = waikiki, aes(x = date, y = tmax, color = name)) + 
  geom_point() + 
  geom_line(data = central_park)
```
# This gives you a color to the name
```{r}
ggplot(weather_df, aes(x = tmin, y = tmax)) + 
  geom_point(aes(color = name), alpha = .5) +
  geom_smooth(se = FALSE)
```

# Multi panel plots
#Faceting gives you a fair distance in creating plots (multiple panels of plots)
```{r}
ggplot(weather_df, aes(x = tmin, y = tmax, color = name)) + 
  geom_point(alpha = .5) +
  geom_smooth(se = FALSE) + 
  facet_grid(. ~ name)

ggp_box = weather_df %>%
  ggplot(aes(x = name, y = tmax)) +
  geom_boxplot()
```
```{r}
devtools::install_github("thomasp85/patchwork") 
1
```
```{r}
tmax_tmin_p = 
  weather_df %>% 
  ggplot(aes(x = tmax, y = tmin, color = name)) + 
  geom_point(alpha = .5) +
  theme(legend.position = "none")

prcp_dens_p = 
  weather_df %>% 
  filter(prcp > 0) %>% 
  ggplot(aes(x = prcp, fill = name)) + 
  geom_density(alpha = .5) + 
  theme(legend.position = "none")

tmax_date_p = 
  weather_df %>% 
  ggplot(aes(x = date, y = tmax, color = name)) + 
  geom_point(alpha = .5) +
  geom_smooth(se = FALSE) + 
  theme(legend.position = "bottom")

(tmax_tmin_p + prcp_dens_p) / tmax_date_p
```

```{r}
weather_df %>%
  mutate(name = forcats::fct_relevel(name, c("Waikiki_HA", "CentralPark_NY", "Waterhole_WA"))) %>% 
  ggplot(aes(x = name, y = tmax)) + 
  geom_violin(aes(fill = name), color = "blue", alpha = .5) + 
  theme(legend.position = "bottom")
```
 #reorder 
```{r}
weather_df %>%
  mutate(
    name = factor(name),
    name = fct_reorder(name, tmax)
  ) %>%
  ggplot(aes(x = name, y = tmax, color = name)) +
  geom_boxplot()
```
 
```{r}
weather_df %>%
  select(name, tmax, tmin) %>% 
  pivot_longer(
    tmax:tmin,
    names_to = "observation", 
    values_to = "temp") %>% 
  ggplot(aes(x = temp, fill = observation)) +
  geom_density(alpha = .5) + 
  facet_grid(~name) + 
  viridis::scale_fill_viridis(discrete = TRUE)
```
 
